---
title: "Data Documentation"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
> Data Generation

All the single molecule imaging data was generated using wide-field 
epi-illumination microscope. Single molecules were tracked using 
DiaTrack v3.05 software. Single molecule images were collected 
after pre-bleach of initial intense fluorescence (glow), and 405-nm laser 
excitation was added as needed to maintain desired fluorophore detection 
density (reference to a manuscript will be added soon which contains detailed 
information on imaging regime setup). 

> Folders of Imaging Data

--- HSF and HSF_2 

Single molecule tracking data of Heat Shot Factor (HSF) in Drosophila, HSF is 
labeled with Halo tag. Data is a courtesy of Xiaona Tang. 

--- HTZ1 and ImageJ

Single molecule tracking data of histone variant HTZ1 in fission yeast, HTZ1 is 
labeled with Halo tag. Data is a courtesy of Anand Ranjan.

--- SWR1 and SWR1_2

Single molecule tracking data of histone variant HTZ1 chromatin remodelers SWR1 
in fission yeast, SWR1 is labeled with Halo tag. Data is a courtesy of 
Anand Ranjan.

Within each folder, 

  + *_MASK.tif*, are mask files for removing background, generated by 
  thresholding images from initial fluorescence (glow).

  + *_Nuclei.tif*, are the Nuclei images for plotting, generated by the DIC 
  image of the cell.

  + *.mat* files are the matlab file outputs from tracking software Diatrack;

  + *.csv* files are the matlab file outputs from Diatrack converted to .csv by 
  first reading the .mat files with `createTrackll()` and exporting to .csv 
  files with `exportTrackll()`. The .csv files from `exportTrackll()` have the 
  same format as output files from ImageJ plugin Particle Tracker. There are 
  also no functional difference between the tracklls created from .mat or .csv 
  files; however, .csv files are used here for their small file sizes and 
  faster reading.

> Other Folders

--- INDEX

  + componentTrackID-SWR1.comp.1.csv

  + componentTrackID-SWR1.comp.1.csv
    
Index files which contains indices of trajectories that are of different 
component, generated by `selComponentTracks()` when output set to be TRUE. It 
used for plotting trajectories of different component. 
    
  + IndexFile.csv

  + IndexFile2.csv

Index file which contains index of interest. The content is compied from first 
two columns of `Dcoef()` function.

---PREF

  + profile.csv
  
Preference file which contains adjustable parameters for `Dcoef()` function. 

```{r}

```
