---
title: "Data Documentation"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
> Data Generation

All the single molecule imaging data was generated using wide-field epi-illumination microscope. Single molecules were tracked using DiaTrack v3.05 software. Single molecule images were collected after pre-bleach of initial intense fluorescence (glow), and 405-nm laser excitation was added as needed to maintain desired fluorophore detection density (reference to a manuscript will be added soon which contains detailed information on imaging regime setup).


> Folders of Imaging Data

--- HSF and HSF_2 

Single molecule tracking data of Heat Shot Factor (HSF) in Drosophila, HSF is labeled with Halo tag. Data is a courtesy of Xiaona Tang. 

--- HTZ1 and ImageJ

Single molecule tracking data of histone variant HTZ1 in fission yeast, HTZ1 is labeled with Halo tag. Data is a courtesy of Anand Ranjan.

--- SWR1 and SWR1_2

Single molecule tracking data of histone variant HTZ1 chromatin remodelers SWR1 in fission yeast, SWR1 is labeled with Halo tag. Data is a courtesy of Anand Ranjan.

Within each folder, 

  + *.._MASK.tif*, is the mask file for removing background, generated by image from initial intense fluorescence (glow).

  + *.._Nuclei.tif*, is the Nuclei image for plotting, generated by the DIC image of the cell.

  + *.mat* file is the matlab file output from tracking software Diatrack;

  + *.txt* file is a txt file output from tracking software Diatrack;

  + *.csv* file is the output file from ImageJ plugin Particle Tracker. Note output file from exportTrackll() function also has the same format.

  

> Other Folders

--- INDEX

  + componentTrackID-SWR1.comp.1.csv

  + componentTrackID-SWR1.comp.1.csv
    
Index file which contains indices of trajectories that are of different component, generated by selComponentTracks function when output set to be TRUE. It is used for plotting trajectories of different component. 
    
  + IndexFile.csv

  + IndexFile2.csv

Index file which contains index of interest. The content is compied from first two columns of Dcoef() function.

---PREF

  + profile.csv
  
Preference file which contains adjustable parameters for Dcoef() function. 

```{r}

```
